---
title: "Introduction to Snakemake"
author: 
  - "Dillon Barker"
date: '2021-10-22'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: false

---

```{r setup, include=FALSE}
library(DiagrammeR)
library(stringr)
library(tufte)
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#1381B0",
  secondary_color = "#FF961C",
  inverse_header_color = "#FFFFFF",
  text_font_size = "1.25rem"
)

```

class: inverse, center, middle
name: lesson3

# Lesson 3
## 2021-10-22

---

## Lesson 3

1. Assignment 2 Answer
2. Fixing when things go wrong
3. Priorities
4. Script files (Python, R, and Julia)
5. Modularizing workflows
6. Visualizing Workflows
7. Linting
8. Clinic

---
## Assignment 2 Answer

### Conda Environment Files
```
snakefiles/
├── asn-2.smk
└── envs/
    ├── pirate.yaml
    └── prokka.yaml
```

.pull-left[
```yaml
# envs/prokka.yaml
name: prokka
channels:
        - conda-forge
        - bioconda
dependencies:
        - prokka
```
]
.pull-right[
```yaml
# envs/pirate.yaml
name: pirate
channels:
        - bioconda
        - conda-forge
dependencies:
        - pirate
```
]
---
## Assignment 2 Answer
### Snakemake Workflow
https://github.com/dorbarker/snakemake-intro/blob/main/answer_keys/asn-2.smk
---

## When Things Go Wrong
### Locked Working Directories

- Snakemake locks its working directory when running
     - Prevents other snakemake instances from running in the same place

- Snakemake removes the lock when it completes (success _or_ failure)

--

- Lockfile may not be removed when if snakemake crashes or is killed by slurm
    - `scancel --user your_username`


```sh
Unable to lock working directory.
```

---
## Unlocking a Stale Lock

#### The Correct Way: `--unlock`

```bash
sbatch -c 1 --mem 2G --wrap=\
"snakemake --unlock -s path/to/your/workflow.smk -d your/locked/analysis/ -j 1"
```
--
#### ☠☠☠ Nuclear Option ☠☠☠

```bash
rm -r your/locked/analysis/.snakemake
```
---
## When Things Go Wrong
### It Should Have Worked But It Didn't!

- Sometimes a job finishes successfully but snakemake doesn't detect the outputs

- Filesystem latency _especially_ on a cluster like Waffles can be a factor
    - It takes some non-zero amount of time to write results to disk

- Consider setting `--latency-wait` with larger number (default: `5` seconds)

---
## Priorities

.pull-left[
- Rule priorities may be set

- These **are not** the same as slurm partitions

- Only determine the priority of execution _within_ the workflow

- Rules default to priority `0`

- Bigger numbers are higher priority
]
--
.pull-right[
```python
rule my_example:
    input: ...
    output: ...
    priority: 50
    shell: ...
```
]

---
## Script Files

- Not to be confused with an external command that happens to be a script

--

- Like a `run` block, but stored in another file
    - Path is specified relative to the workflow like conda envs

--

- Available languages:
    - Python
    - R
    - Julia

--

- Snakemake will use special variables in the script
    - These scripts won't be general purpose!
    - Tied to snakemake

---
## Script Files

```python
# workflow.smk
rule scriptfile_example:
    input:
        "path/to/inputfile",
        "path/to/other/inputfile"
    output:
        "path/to/outputfile",
        "path/to/another/outputfile"
    script:
        "scripts/script.py"
```

```python
# scripts/script.py
def do_something(data_path, out_path, threads, myparam):
    # python code here

do_something(snakemake.input[0], snakemake.output[0], 
             snakemake.threads, snakemake.config["myparam"])
```
---
## Including Workflows

- Existing workflows can be included with `include`

- Allows more modular design of workflows

- Like copy/pasting another workflow in

--

```python
rule all:
    input: "summary_report.txt"

include: "summary_report.smk"
include: "determine_amr.smk"
include: "assemble_genomes.smk"
```
---
## Visualizing Workflows


```bash
# Shows just the abstract graph of rules
snakemake -s your/workflow.smk --rulegraph | dot -Tpng > workflow.png
# OR
# Show every sample's journey through the workflow
snakemake -s your/workflow.smk --dag | dot -Tpng > workflow.png
```


```{r, snakemake-workflow, echo=FALSE, fig.height=3.5}
workflow <- paste0(readLines("figures/snakemake-rule-graph.dot"))
grViz(workflow)

```
---
## Linting
> **Lint**, or a **linter**, is a static code analysis tool 
> used to flag programming errors, bugs, stylistic 
> errors and suspicious constructs.
>
> `r tufte::quote_footer('— Wikipedia')`

--

- Get automated _suggestions_ for your workflow!

--

```bash
snakemake --lint -s path/toyour/workflow.smk
```
---
## Colophon

- This presentation is written in RMarkdown + Xaringan
    - Built with Snakemake

### GitHub Repository
https://github.com/dorbarker/snakemake-intro/

### Today's Lecture
https://github.com/dorbarker/snakemake-intro/blob/main/lessons/snakemake-intro-lesson-3.pdf

---
class: center, middle

# Clinic
## Please feel free to ask any questions!