---
title: "Introduction to Snakemake"
author: 
  - "Dillon Barker"
date: '2021-10-22'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: false

---
  
```{r setup, include=FALSE}
library(DiagrammeR)
library(stringr)
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#1381B0",
  secondary_color = "#FF961C",
  inverse_header_color = "#FFFFFF",
  text_font_size = "1.25rem"
)

```
class: inverse, center, middle
name: lesson3

# Lesson 3
## 2021-10-22

---

## Lesson 3

1. Assignment 2 Answer
2. Fixing when things go wrong
3. Priorities
4. Script files (Python, R, and Julia)
5. Modularizing workflows
6. Visualizing Workflows
7. Clinic

---

## Assignment 2
  
---

## When Things Go Wrong
### Locked Working Directories

- Snakemake locks its working directory when running
     - Prevents other snakemake instances from running in the same place

- Snakemake removes the lock when it completes (success _or_ failure)

- Lockfile may not be removed when if snakemake crashes or is killed by slurm
    - `scancel --user your_username`


```sh
Unable to lock working directory.
```

---
  
## Unlocking a Stale Lock
  
#### The Correct Way: `--unlock`
  
```bash
sbatch -c 1 --mem 2G --wrap "snakemake --unlock -s path/to/your/workflow.smk -j 1"
```

#### ☠☠☠ Nuclear Option ☠☠☠

```bash
rm -r ./.snakemake
```
---
## When Things Go Wrong
### It Should Have Worked But It Didn't!

- Sometimes a job finishes successfully but snakemake doesn't detect the outputs

- Filesystem latency _especially_ on a cluster like Waffles can be a factor
    - It takes some non-zero amount of time to write results to disk

- Consider setting `--latency-wait` with larger number (default: `5` seconds)

---
## Priorities

.pull-left[
- Rule priorities may be set

- These **are not** the same as slurm partitions

- Only determine the priority of execution _within_ the workflow

- Rules default to priority `0`

- Bigger numbers are higher priority
]

.pull-right[
```python
rule my_example:
    input: ...
    output: ...
    priority: 50
    shell: ...
```
]

---
## Script Files

- Not to be confuse with an external command 

- Like a `run` block, but stored in another file
    - Path is specified relative to the workflow

- Available languages:
    - Python
    - R
    - Julia
    
- Snakemake will use special variables in the script
    - These scripts won't be general purpose!
    - Tied to snakemake

---
## Script Files

```python
# workflow.smk
rule scriptfile_example:
    input:
        "path/to/inputfile",
        "path/to/other/inputfile"
    output:
        "path/to/outputfile",
        "path/to/another/outputfile"
    script:
        "scripts/script.py"
```

```python
# scripts/script.py
def do_something(data_path, out_path, threads, myparam):
    # python code here

do_something(snakemake.input[0], snakemake.output[0], 
             snakemake.threads, snakemake.config["myparam"])
```
---
## Including Workflows

- Existing workflows can be included with `include`

- Allows more modular design of workflows

- Like copy/pasting another workflow in

```python
rule all:
    input: "summary_report.txt"

include: "summary_report.smk"
include: "determine_amr.smk"
include: "assemble_genomes.smk"
```
---
## Visualizing Workflows


```bash
# Shows just the abstract graph of rules
snakemake -s your/workflow.smk --rulegraph | dot -Tpng > workflow.png
# OR
# Show every sample's journey through the workflow
snakemake -s your/workflow.smk --dag | dot -Tpng > workflow.png
```


```{r, snakemake-workflow, echo=FALSE, fig.height=3.5}
workflow <- paste0(readLines("figures/snakemake-rule-graph.dot"))
grViz(workflow)

```

---
class: center, middle

# Clinic
## Please feel free to ask any questions!